### 2020-03-26 manuscrit.tex

%% Use the standard UP-methodology class
%% with French language.
%%
%% You may specify the option 'twoside' or 'oneside' for
%% the document.
%%
%% See the documentation tex-upmethodology on
%% http://www.arakhne.org/tex-upmethodology/
%% for details about the macros that are provided by the class and
%% to obtain the list of the packages that are already included.
\pdfobjcompresslevel 0
\documentclass[english,standardlists]{spimubphdthesis}

%% Custom commands and packages configuration
\input{texcmds}

%\loadglsentries[main]{glossaire}
\loadglsentries[\acronymtype]{resources/acronyms}

%%--------------------
%% Set the title, subtitle, defense date, and
%% the registration number of the PhD thesis.
%% The optional parameter is the subtitle of the PhD thesis.
%% The first mandatory parameter is the title of the PhD thesis.
%% The second mandatory parameter is the date of the PhD defense.
%% The third mandatory parameter is the location/city of the PhD defense.
%% The forth mandatory parameter is the reference number given by
%% the University Library after the PhD defense.
\declarethesis[]{Titre de la thèse}{Date de soutenance}{Dijon}{XXX}

%%--------------------
%% Set the author of the PhD thesis
\addauthor[romain.cendre@gmail.com]{Romain}{CENDRE}

%%--------------------
%% Add a member of the jury
%%
%% CAUTION 1: If a Jury member is not present during the defense,
%%            she/he must be in the list of the Jury members.
%%            Only the reviewers and the members who are present during the defense must
%%            appear in the Jyry member list.
%% CAUTION 2: After your defense, you must assign the role "Pr\'esident" to
%%            the Jury member who have been the President of the Jury.
%% CAUTION 3: The recommended order for the Jury members is:
%%            President, Reviewer(s), Examiner(s), Director(s),
%%            Other supervisor(s), Invited person(s).
%% \addjury{Firstname}{Lastname}{Role in the jury}{Position}
\addjury{Prenom}{Nom}{Rapporteur}{Professeur, Université XXX}
\addjury{Prenom}{Nom}{Rapporteur}{Professeur, Université XXX}
\addjury{Prenom}{Nom}{Directeur de thèse}{Professeur, Université XXX}
\addjury{Prenom}{Nom}{Co-encadrant}{Maître de Conférences, Université XXX}

%%--------------------
%% Change style of the table of the jury
%% \Set{jurystyle}{put macros for the style}
%\Set{jurystyle}{\small}

%%--------------------
%% Set the English abstract
\thesisabstract[english]{
English resume
}

%%--------------------
%% Set the English keywords. They only appear if
%% there is an English abstract
\thesiskeywords[english]{english,keywords}

%%--------------------
%% Set the French abstract
\thesisabstract[french]{
Résumé français
}

%%--------------------
%% Set the French keywords. They only appear if
%% there is an French abstract
\thesiskeywords[french]{mots clés,example}

%%--------------------
%% Change the layout and the style of the text of the "primary" abstract.
%% If your document is written in French, the primary abstract is in French,
%% otherwise it is in English.
%\Set{primaryabstractstyle}{\tiny}

%%--------------------
%% Change the layout and the style of the text of the "secondary" abstract.
%% If your document is written in French, the secondary abstract is in English,
%% otherwise it is in French.
%\Set{secondaryabstractstyle}{\tiny}

%%--------------------
%% Change the layout and the style of the text of the "primary" keywords.
%% If your document is written in French, the primary keywords are in French,
%% otherwise they are in English.
%\Set{primarykeywordstyle}{\tiny}

%%--------------------
%% Change the layout and the style of the text of the "secondary" keywords.
%% If your document is written in French, the secondary keywords are in English,
%% otherwise they are in French.
%\Set{secondarykeywordstyle}{\tiny}

%%--------------------
%% Change the speciality of the PhD thesis
\Set{speciality}{Informatique}

%%--------------------
%% Change the institution
\Set{universityname}{Université de Bourgogne}

%%--------------------
%% Clear the list of the laboratories
\resetlaboratories

%%--------------------
%% Add the laboratory where the thesis was made
%\addlaboratory{Laboratoire Waynes Industry}
\addlaboratory{ImVia}

%%--------------------
%% Clear the list of the partner/sponsor logos
\resetpartners

%%--------------------
%% Add the logos of the partners or the sponsors on the front page
%%
%% CAUTION 1: At least, the logo of the University should appear (UB)
%%
%\addpartner[image options]{image name}

\addpartner{resources/logos/ub}

%%--------------------
%% Change the header and the foot of the pages.
%% You must include the package "fancyhdr" to
%% have access to these macros.
%% Left header
%\lhead{}
%% Center header
%\chead{}
%% Right header
%\rhead{}
%% Left footer
%\lfoot{}
%% Center footer
%\cfoot{}
%% Right footer
%\rfoot{}

%%--------------------
% Declare several theorems
%\declareupmtheorem{mytheorem}{My Theorem}{List of my Theorems}

%%--------------------
%% Change the message on the backcover.
\Set{backcovermessage}{
  % Ajouter logos partenaires de la thèse
}

% fancy quotes
\definecolor{quotemark}{gray}{0.7}
\makeatletter
\def\fquote{%
    \@ifnextchar[{\fquote@i}{\fquote@i[]}%]
           }%
\def\fquote@i[#1]{%
    \def\tempa{#1}%
    \@ifnextchar[{\fquote@ii}{\fquote@ii[]}%]
                 }%
\def\fquote@ii[#1]{%
    \def\tempb{#1}%
    \@ifnextchar[{\fquote@iii}{\fquote@iii[]}%]
                      }%
\def\fquote@iii[#1]{%
    \def\tempc{#1}%
    \vspace{1em}%
    \noindent%
    \begin{list}{}{%
         \setlength{\leftmargin}{0.55\textwidth}%
         \setlength{\rightmargin}{0.01\textwidth}%
                  }%
         \item[]%
         \begin{picture}(0,0)%
         \put(-15,-5){\makebox(0,0){\scalebox{3}{\textcolor{quotemark}{``}}}}%
         \end{picture}%
         \begingroup\itshape}%
 %%%%********************************************************************
 \def\endfquote{%
 \endgroup\par%
 \makebox[0pt][l]{%
 \hspace{0.4\textwidth}%
 \begin{picture}(0,0)(0,0)% 
	 \put(15,15){\makebox(0,0){%
	 \scalebox{3}{\color{quotemark}''}}}%
 \end{picture}}%
 \ifx\tempa\empty%
 \else%
    \ifx\tempc\empty%  
       \hfill\rule{100pt}{0.5pt}\\\mbox{}\hfill\tempa,\ \emph{\tempb}%
   \else%		
       \hfill\rule{100pt}{0.5pt}\\\mbox{}\hfill\tempa,\ \emph{\tempb},\ \tempc%
   \fi\fi\par%
   \vspace{0.5em}%
 \end{list}%
 }%
\makeatother

\begin{document}
\dominitoc

%%--------------------
%% The following line does nothing until
%% the class option 'nofrontmatter' is given.
%\frontmatter

%%--------------------
%% The following line permits to add a chapter for "acknowledgements"
%% at the beginning of the document. This chapter has not a chapter
%% number (using the "star-ed" version of \chapter) to prevent it to
%% be in the table of contents

	\parbox[t]{\textwidth}{
	\vspace{1.5cm}
		\begin{fquote}[Bernard de Chartres][\textit{Gloses sur Priscien}]
			Des nains sur des épaules de géants.
		\end{fquote}	
	}

\chapter*{Remerciements}
Ces travaux ont été rendus possibles par le Conseil Régional de Bourgogne Franche Comté (CBR) et le Fonds européen de développement régional (FEDER) dans le cadre de l'initiative des Jeunes Chercheurs Entrepreneurs (JCE).\par
Je remercie tout naturellement ma structure d'accueil, le laboratoire Imagerie et Vision Artificielle (ImViA) et l'ensemble de ses membres permanents et non permanents qui m'auront permis de travailler dans un cadre des plus favorables durant ces 3 années. Je remercie fortement Franck Marzani ainsi que Alamin Mansouri pour leur précieux soutien moral autant que pour leur oreille attentive et ce malgré les nombreuses responsabilités qui les incombent.\par
- Médecins
- Collègues
- Proches
- Région

% Exmeple Pierre
%  Je remercie les enseignants chercheurs et autres permanents du laboratoire pour leur contribution non négligeable à l'atmosphère chaleureuse dans laquelle j'ai pu évoluer, en particulier madame Fan Yang Song et messieurs Yannick Benezeth et Barthélémy Heyrman. \\		
% Je remercie les membres de l'équipe technique, messieurs Cédric Clerc et Romain Troisgros pour le temps qu'ils ont su me consacrer, pour leur grande aide et leur précieux conseils sur l'élaboration du prototype et la conception des pièces. \`A monsieur Matthieu Rossé pour la conception des boitiers des Smart Camera Rasperry. \`A monsieur Vincent Thivent et sa société ODALID pour le prêt de son lecteur de cartes RFID. J'aimerai également exprimer ma gratitude envers monsieur David Darson, ingénieur de recherche au laboratoire de physique de l'\'Ecole Normale Supérieure de Paris (LPENS), pour sa disponibilité, son expertise et ses conseils qui m'ont été d'une grande aide dans la conception du prototype de scanner rotatif présenté dans cette thèse.  \par 
% Je souhaite remercier chaleureusement mes collègues et amis doctorants, pour avoir partagé ces trois années de thèse et sans qui elles n'auraient pas eu la même saveur. Je pense notamment à mesdames Yangjin Bi, Margarita Khokhlova et à messieurs Serge Bobbia, Yoan Marin, Anthony Boscaro, Axel Moinet, Nicolas Brochard, Richard Macwan, Roberto Marroquin, Romain Cendre. \par
% Bien évidemment, je remercie mes parents pour m'avoir soutenu dans mes projets et mon parcours, et de m'avoir permis d'arriver jusque-là. Pour terminer, je tiens à exprimer toute ma reconnaissance à ma compagne, Audrey Febvre, pour avoir su me motiver et me supporter lors de cette thèse. Son soutien en toutes circonstances et ses encouragements m'ont accompagné sans interruption et ont sans aucun doute joué un rôle majeur dans l'aboutissement de cette thèse. Je dédie le fruit de ces travaux à mes parents, tout juste gradés grands-parents, à ma compagne qui est déjà une parfaite maman et à mon fils qui fait le bonheur de ses jeunes parents.

%%--------------------
%% Include a general table of contents
%\phdseetodo
%\phdtodotoc

\tableofcontents

%%--------------------
%% The content of the PhD thesis
\mainmatter

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=0.9\linewidth]{contents/introduction/resources/evolutiondermatologists.pdf}
%     \caption{Evolution du nombre de dermatologues en France.}
%     \label{fig:number}
% \end{figure}\par
% Et voilà \Cref{fig:number}
% Inclure les fichiers contenant les différents chapitres
% Introduction
\include{contents/x_introduction/content}

% Part 1
\part{Contexte}
\label{part:contexte}
\include{contents/chapter_1/content}
\include{contents/chapter_2/content}
\include{contents/chapter_3/content}
% Part 2
\part{Microscopie confocale par réflectance}
\label{part:microscopie}
\include{contents/x_microscopy/content}
\include{contents/chapter_4/content}
\include{contents/chapter_5/content}
\include{contents/chapter_6/content}

% Part 3
\part{Multimodalité}
\label{part:multimodal}
\include{contents/x_multimodal/content}
\include{contents/chapter_7/content}

% Conclusion
\include{contents/x_conclusion/content}

%%--------------------
%% Start the end of the thesis
\backmatter

%%--------------------
%% Bibliography
\bibliographystyle{plain}
\bibliography{resources/bibliography}

%%--------------------
%% List of figures and tables

%% Include a chapter with a list of all the figures.
%% In French typograhic standard, this list must be at
%% the end of the document.
\listoffigures

%% Include a chapter with a list of all the tables.
%% In French typograhic standard, this list must be at
%% the end of the document.
\listoftables

%%--------------------
%% Include a list of definitions
%\listofdefinitions

%% --------------------
%% Include a list of acronyms and a glossary (fichier acronymes.tex)
%\glsaddall
\printglossary[type=\acronymtype,style=listfracr,title={Liste des acronymes}]
%% Si besoin d'un glossaire, le remplir dans glossaire.tex et décommenter 
%\printglossary

%%--------------------
%% Appendixes
\appendix
\input{contents/annexes.tex}

\end{document}