### 2020-04-26 manuscrit.tex

%% Use the standard UP-methodology class
%% with French language.
%%
%% You may specify the option 'twoside' or 'oneside' for
%% the document.
%%
%% See the documentation tex-upmethodology on
%% http://www.arakhne.org/tex-upmethodology/
%% for details about the macros that are provided by the class and
%% to obtain the list of the packages that are already included.
\pdfobjcompresslevel 0
\documentclass[english,standardlists]{spimubphdthesis}

% To prevent warning on todo package
\setlength{\marginparwidth}{2cm}

%% Custom commands and packages configuration
\input{texcmds}

%\loadglsentries[main]{glossaire}
\loadglsentries[\acronymtype]{resources/acronyms}

%%--------------------
%% Set the title, subtitle, defense date, and
%% the registration number of the PhD thesis.
%% The optional parameter is the subtitle of the PhD thesis.
%% The first mandatory parameter is the title of the PhD thesis.
%% The second mandatory parameter is the date of the PhD defense.
%% The third mandatory parameter is the location/city of the PhD defense.
%% The forth mandatory parameter is the reference number given by
%% the University Library after the PhD defense.
\declarethesis[]{Titre de la thèse}{Date de soutenance}{Dijon}{XXX}

%%--------------------
%% Set the author of the PhD thesis
\addauthor[romain.cendre@gmail.com]{Romain}{CENDRE}

%%--------------------
%% Add a member of the jury
%%
%% CAUTION 1: If a Jury member is not present during the defense,
%%            she/he must be in the list of the Jury members.
%%            Only the reviewers and the members who are present during the defense must
%%            appear in the Jyry member list.
%% CAUTION 2: After your defense, you must assign the role "Pr\'esident" to
%%            the Jury member who have been the President of the Jury.
%% CAUTION 3: The recommended order for the Jury members is:
%%            President, Reviewer(s), Examiner(s), Director(s),
%%            Other supervisor(s), Invited person(s).
%% \addjury{Firstname}{Lastname}{Role in the jury}{Position}
\addjury{Prenom}{Nom}{Rapporteur}{Professeur, Université XXX}
\addjury{Prenom}{Nom}{Rapporteur}{Professeur, Université XXX}
\addjury{Prenom}{Nom}{Directeur de thèse}{Professeur, Université XXX}
\addjury{Prenom}{Nom}{Co-encadrant}{Maître de Conférences, Université XXX}

%%--------------------
%% Change style of the table of the jury
%% \Set{jurystyle}{put macros for the style}
%\Set{jurystyle}{\small}

%%--------------------
%% Set the English abstract
\thesisabstract[english]{
English resume
}

%%--------------------
%% Set the English keywords. They only appear if
%% there is an English abstract
\thesiskeywords[english]{Dermatology, Lentigo, Machine Learning, Deep Learning, Image processing, Image classification}

%%--------------------
%% Set the French abstract
\thesisabstract[french]{
Résumé français
}

%%--------------------
%% Set the French keywords. They only appear if
%% there is an French abstract
\thesiskeywords[french]{Dermatologie, Lentigo, Apprentissage automatique, Apprentissage profond, Traitement d'image, Classification d'images}

%%--------------------
%% Change the layout and the style of the text of the "primary" abstract.
%% If your document is written in French, the primary abstract is in French,
%% otherwise it is in English.
%\Set{primaryabstractstyle}{\tiny}

%%--------------------
%% Change the layout and the style of the text of the "secondary" abstract.
%% If your document is written in French, the secondary abstract is in English,
%% otherwise it is in French.
%\Set{secondaryabstractstyle}{\tiny}

%%--------------------
%% Change the layout and the style of the text of the "primary" keywords.
%% If your document is written in French, the primary keywords are in French,
%% otherwise they are in English.
%\Set{primarykeywordstyle}{\tiny}

%%--------------------
%% Change the layout and the style of the text of the "secondary" keywords.
%% If your document is written in French, the secondary keywords are in English,
%% otherwise they are in French.
%\Set{secondarykeywordstyle}{\tiny}

%%--------------------
%% Change the speciality of the PhD thesis
\Set{speciality}{Informatique}

%%--------------------
%% Change the institution
\Set{universityname}{Université de Bourgogne}

%%--------------------
%% Clear the list of the laboratories
\resetlaboratories

%%--------------------
%% Add the laboratory where the thesis was made
%\addlaboratory{Laboratoire Waynes Industry}
\addlaboratory{ImVia}

%%--------------------
%% Clear the list of the partner/sponsor logos
\resetpartners

%%--------------------
%% Add the logos of the partners or the sponsors on the front page
%%
%% CAUTION 1: At least, the logo of the University should appear (UB)
%%
%\addpartner[image options]{image name}

\addpartner{resources/logos/ub}

%%--------------------
%% Change the header and the foot of the pages.
%% You must include the package "fancyhdr" to
%% have access to these macros.
%% Left header
%\lhead{}
%% Center header
%\chead{}
%% Right header
%\rhead{}
%% Left footer
%\lfoot{}
%% Center footer
%\cfoot{}
%% Right footer
%\rfoot{}

%%--------------------
% Declare several theorems
%\declareupmtheorem{mytheorem}{My Theorem}{List of my Theorems}

%%--------------------
%% Change the message on the backcover.
\Set{backcovermessage}{
  % Ajouter logos partenaires de la thèse
}

% fancy quotes
\definecolor{quotemark}{gray}{0.7}
\makeatletter
\def\fquote{%
    \@ifnextchar[{\fquote@i}{\fquote@i[]}%]
           }%
\def\fquote@i[#1]{%
    \def\tempa{#1}%
    \@ifnextchar[{\fquote@ii}{\fquote@ii[]}%]
                 }%
\def\fquote@ii[#1]{%
    \def\tempb{#1}%
    \@ifnextchar[{\fquote@iii}{\fquote@iii[]}%]
                      }%
\def\fquote@iii[#1]{%
    \def\tempc{#1}%
    \vspace{1em}%
    \noindent%
    \begin{list}{}{%
         \setlength{\leftmargin}{0.55\textwidth}%
         \setlength{\rightmargin}{0.01\textwidth}%
                  }%
         \item[]%
         \begin{picture}(0,0)%
         \put(-15,-5){\makebox(0,0){\scalebox{3}{\textcolor{quotemark}{``}}}}%
         \end{picture}%
         \begingroup\itshape}%
 %%%%********************************************************************
 \def\endfquote{%
 \endgroup\par%
 \makebox[0pt][l]{%
 \hspace{0.4\textwidth}%
 \begin{picture}(0,0)(0,0)% 
	 \put(15,15){\makebox(0,0){%
	 \scalebox{3}{\color{quotemark}''}}}%
 \end{picture}}%
 \ifx\tempa\empty%
 \else%
    \ifx\tempc\empty%  
       \hfill\rule{100pt}{0.5pt}\\\mbox{}\hfill\tempa,\ \emph{\tempb}%
   \else%		
       \hfill\rule{100pt}{0.5pt}\\\mbox{}\hfill\tempa,\ \emph{\tempb},\ \tempc%
   \fi\fi\par%
   \vspace{0.5em}%
 \end{list}%
 }%
\makeatother

\begin{document}
\dominitoc

%%--------------------
%% The following line does nothing until
%% the class option 'nofrontmatter' is given.
%\frontmatter

%%--------------------
%% The following line permits to add a chapter for "acknowledgements"
%% at the beginning of the document. This chapter has not a chapter
%% number (using the "star-ed" version of \chapter) to prevent it to
%% be in the table of contents

	\parbox[t]{\textwidth}{
	\vspace{1.5cm}
		\begin{fquote}[Bernard de Chartres][\textit{Gloses sur Priscien}]
			Des nains sur des épaules de géants.
		\end{fquote}	
	}

\chapter*{Remerciements}
Ces travaux ont été rendus possibles par le Conseil Régional de Bourgogne Franche Comté et le \gls{feder} dans le cadre de l'initiative des \gls{jce}. Je tiens naturellement à leur exprimer mes plus sincères remerciements, sans quoi toute cette aventure n'aurait pas pu voir le jour. En effet, ces trois années auront été l'occasion d'une expérience scientifique intense, mais également de cours prenant auprès de l'IAE de Dijon et de la Business School of Burgundy.\par

A ce titre, je remercie fortement Franck Marzani ainsi que Alamin Mansouri d'avoir considéré et soutenu ma candidature à ce financement et à ce projet commun. Par ailleurs, je les remercient également pour leur précieux soutien moral mais également pour leur oreille attentive et bienveillante et ce malgré les nombreuses responsabilités qui les incombent. Je dédie, par la même occasion ces propos à deux médecins, Jean-Luc Perrot et Elisa Cinotti, dont la participation, le temps consacré, ainsi que la bonne humeur auront été essentiel à la réussite de ce projet.\par

Cette page est également le lieu pour exprimer ma gratitude envers mon environnement proche professionnel. Ainsi, je remercie tout naturellement ma structure d'accueil, le laboratoire \gls{imvia} et l'ensemble de ses membres permanents (administratif et non administratif) et non permanents qui m'auront permis de travailler dans un cadre des plus favorables durant ces 3 années. Je pense essentiellement à Fabienne Blaise, Anne Fournier, et Khadija Jourani qui m'ont épaulé pour certaines démarches administratives, et à Cyrille Migniot, Barthelemy Heyrman, et Julien Dubois qui ont beaucoup contribué pour rendre la vie au quotidien plus aisée.\par

J'ai une pensée toute particulière pour l'ensemble de mes collègues de laboratoire et de Master, qui par leur présence m'ont rendu ces trois années plus conviviale. Ne pouvant leur consacrer une pensée particulière faute de place, je me résout à des listes. Au sein du laboratoire, je pense à: Abdelali Douiyek, Mustapha Bouderbane, Axel Moinet, Margarita Khokhlova, Yoan Marin, Pierre Bonazza, Duncan Luguern. En ce qui concerne le master: Jérémie Vion et Antoine Greisner auront été deux super collègues de travaux, mais également des collègues de \gls{jce} dont Alicia Froidurot, Garance Dejouy et Anne Berrou ont fait partie. Enfin, je profite de ce paragraphe pour dresser une liste toute particulière à notre noyau de course soudé composé de: Fan Yang, Yannick Benezeth, Alexandre Krebs, Abir Zendagui et Rita Meziati.\par

Bien évidemment, je terminerait par mes proches, qui ont été constamment à mes côtés durant les moments les plus difficiles et ont eus des paroles réconfortantes quand je le nécessitait. Tout d'abord, mes amis avec qui il n'a pas été facile toujours de garder contact malgré la distance, dont j'aurais vu pour la plupart l'évolution et devenir une famille. Et enfin, mon cercle plus proche, incluant mes parents qui m'ont porté durant toutes ces années et encouragé à aller de l'avant, et depuis 5 années ma compagne, Claire, qui me suis au quotidien et dont la présence me permet de surmonter certaines périodes difficiles.\par

%%--------------------
%% Include a general table of contents
%\phdseetodo
%\phdtodotoc
\tableofcontents

%%--------------------
%% The content of the PhD thesis
\mainmatter

% Introduction
\include{contents/x_introduction/content}

% Part 1
\part{Contexte}
\label{part:contexte}
\include{contents/chapter_1/content}
\include{contents/chapter_2/content}
\include{contents/chapter_3/content}
% Part 2
\include{contents/x_microscopy/content}
\part{Microscopie confocale par réflectance}
\label{part:microscopie}
\include{contents/chapter_4/content}
\include{contents/chapter_5/content}
\include{contents/chapter_6/content}

% Part 3
\include{contents/x_multimodal/content}
\part{Multimodalité}
\label{part:multimodal}
\include{contents/chapter_7/content}

% Conclusion
\include{contents/x_conclusion/content}
\include{contents/x_valorisation/content}

%%--------------------
%% Start the end of the thesis
\backmatter

%%--------------------
%% Bibliography
\bibliographystyle{plain}
\bibliography{resources/bibliography}

%%--------------------
%% List of figures and tables

%% Include a chapter with a list of all the figures (French).
\listoffigures

%% Include a chapter with a list of all the tables (French).
\listoftables

%%--------------------
%% Include a list of definitions
%\listofdefinitions

%% --------------------
%% Include a list of acronyms and a glossary (fichier acronymes.tex)
%\glsaddall
\printglossary[type=\acronymtype,style=listfracr,title={Liste des acronymes}]
%% Si besoin d'un glossaire, le remplir dans glossaire.tex et décommenter 
%\printglossary

%%--------------------
%% Appendixes
\appendix
\input{contents/annexes.tex}

\end{document}